Nonlinear Dendritic Coincidence Detection for Supervised Learning

### 2021-04-19 notes/draft_article.tex
.

		In a supervised learning scheme, where the top down input
		arriving at the apical compartment acts as the teaching signal,
		the most straight-forward learning rule for the basal synaptic
		weights would be derived from an appropriate loss function,
		such as a mean square error, based on the difference between 
		basal and apical input, i.e. $I_{\rm p} - I_{\rm d}$. Theoretical work has 
		investigated possible learning mechanisms
		that could utilize an intracellular error signal
		\citep{Urbanczik_2014,Schiess_2016,Guerguiev_2017}.
		However, a clear experimental
		evidence for a physical quantity encoding such an error 
		is---to our knowledge---yet to be found. 
		On the other hand, Hebbian-type plasticity is extensively
		documented in experiments. Therefore, our work was
		based on the question whether the non-linear interactions
		between basal and apical synaptic input could, when
		combined with a Hebbian plasticity rule, allow a neuron
		to learn to reproduce an apical teaching signal in its
		proximal input.

		In our work, we combined a phenomenological model that generates the output
		firing rate as a function of two streams of synaptic input (subsuming basal and apical input)
		with Hebbian as well as BCM-like plasticity rules on basal synapses. We hypothesized that 
		this combination of neural activation and plasticity rules would lead to an
		increased correlation between basal and apical input.

		Furthermore, this temporal alignment could potentially facilitate apical inputs to act
		as top-down teaching signals, without the need for an explicit error-driven
		learning rule. Thus, we also tested our model in a simple linear 
		supervised classification task and compared it to the performance of a simple
		point neuron equipped with similar plasticity rules.

@@

		In all our numerical experiments, we compared this model with a 
		simple point neuron model that was given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			y(t) = \sigma\left(I_{\rm p}(t) + I_{\rm d}(t) - \theta \right) \; .
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		In all our numerical experiments, the apical input $I_{\rm d}$ was generated
		``as is", meaning, it was not dynamically calculated as a superposition
		of multiple presynaptic inputs, but given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			I_{\rm d}(t) = n_d(t) x_d(t) - b_d(t) \; ,
			\label{eq:I_d}
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $n_d(t)$ is a scaling factor, $x_d(t)$ a pre-generated
		discrete time sequence and $b_d(t)$ a bias. Note that $n_d$ and $b_d$ 
		are time dependent since they were subject to adaptation processes 
		that are described in the next section.

		Similarly, $I_{\rm p}(t)$ was given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			I_{\rm p}(t) = n_p(t) \sum_{i=1}^{N} x_{p,i}(t) w_i(t) - b_p(t) \; ,
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $N$ is the number of presynaptic afferents, $x_{p,i}(t)$ the
		corresponding sequences and $w_i(t)$ the synaptic efficacies.
		As for $I_{\rm d}(t)$, $n_p(t)$ and $b_p(t)$ was used as a time dependent
		scaling and bias.

@@

		We implemented a Hebbian plasticity rule for the basal 
		synaptic weights given by the following update equation:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			w_i(t+1) &= w_i(t) + \mu_w \left(x_{p,i}(t) - \tilde{x}_{p,i}(t)\right)
			\left(y(t) - \tilde{y} \right)
			\label{eq:bcm_plast} \\
			\tilde{x}_{p,i}(t+1) &= (1-\mu_{av})\tilde{x}_{p,i}(t) + \mu_{av}x_{p,i}(t+1) \\
			\tilde{y}(t+1) &= (1-\mu_{av})\tilde{y}(t) + \mu_{av}y(t+1)
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Additionally, we used a
		synaptic normalization constraint
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			w_i(t) \rightarrow \frac{w_i(t)}{\lVert \mathbf{w}(t)\rVert}
			\label{eq:weight_norm}
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		in each time step, where $\lVert \mathbf{w}(t)\rVert$ denotes the Euclidean
		norm of the synaptic weight vector. 
%		The threshold $\theta_y$ was
%		chosen such that LTP is present if neural activity is in the
%		high-activity regime (yellow area in Fig.~\ref{fig:comp_model}), while
%		LTD is present in the low-activity regime 
%		(given by $\alpha$, see Fig.~\ref{fig:comp_model}). Hence, we set
%		$\theta_y = (1+\alpha)/2$.

		For comparative reasons, the point neuron model was equipped with the
		same plasticity rule for the proximal weights as \eqref{eq:bcm_plast}. 
%		except that the threshold $\theta_y$ was dynamically adjusted as
%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		\begin{equation}
%			\theta_y(t+1) = \left(1-\mu_{av}\right)\theta_y(t) + \mu_{av} y^2(t) \; , 
%		\end{equation}
%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		which is an implementation of a running average of the square activity,
%		as used in the classic BCM-rule.

		Additionally, the scaling and bias variables were changing dynamically
		according to the following homeostatic plasticity rules:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			b_p(t+1) &= b_p(t) + \mu_b \left[I_{\rm p}(t) - I_{\rm p}^t\right] \\
			b_d(t+1) &= b_d(t) + \mu_b \left[I_{\rm d}(t) - I_{\rm d}^t\right] \\
			n_p(t+1) &= n_p(t) + \mu_n \left[V_p^t - \left( I_{\rm p}(t) - \tilde{I}_p(t)\right)^2\right] \\
			n_d(t+1) &= n_d(t) + \mu_n \left[V_d^t - \left( I_{\rm d}(t) - \tilde{I}_d(t)\right)^2\right] \\
			\tilde{I}_p(t+1) &= (1-\mu_{av})\tilde{I}_p(t) + \mu_{av} I_{\rm p}(t+1) \\
			\tilde{I}_d(t+1) &= (1-\mu_{av})\tilde{I}_d(t) + \mu_{av} I_{\rm d}(t+1)
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $I_{\rm p}^t$, $I_{\rm d}^t$, $V_p^t$ and $V_p^t$ define targets for the 
		temporal means and variances of $I_{\rm p}$ and $I_{\rm d}$. The dynamic variables 
		$\tilde{I}_p$ and $\tilde{I}_d$ are simply low-pass filtered running 
		averages of $I_{\rm p}$ and $I_{\rm d}$.

@@

		After both $x_{p,i}(t)$ and $x_d(t)$ were generated,
		a simulation was run using all previously described plasticity
		mechanisms until the dynamic variables reached a stationary state.
		After this learning phase, another set of input sequences was generated
		using the previously described protocol and $\rho[I_{\rm p},I_{\rm d}]$ was calculated.
		Note that plasticity was turned off in this phase. This entire procedure
		allowed us to calculate $\rho[I_{\rm p},I_{\rm d}]$ as a function of the distraction
		parameters $s$ and $N_{dist}$. The results for both neuron models 
		is shown in Fig.~\ref{fig:corr_dimension_scaling}. Here, the total number of 
		basal inputs was $N=10$. One can observe a decorrelation transition for 
		both models. However, the compartment model supports a significantly stronger
		distraction in terms of the scaling factor $s$ as compared to the point model.
		This was a first confirmation of our hypothesis that nonlinear interactions
		between basal and apical input could improve learning guided by top-down signals.

@@

		Next, we investigated if the observed differences would also improve
		the performance in an actual supervised learning task.
		For this purpose, we constructed presynaptic basal input as illustrated
		in Fig.~\ref{fig:illustration_classification}.
		Written in vector form,
		each sample from the basal input was generated from the following expression:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			\mathbf{x}_p(t) = \mathbf{b} + \mathbf{a}\left(c(t) + \sigma_a \zeta_a(t) \right) 
			+ s \cdot \sum_{i=1}^{N_{dist}} \zeta_{dist,i}(t) \mathbf{v}_{dist,i} \; .
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $\mathbf{b}$ is a random vector drawn uniformly from
		$(0,1)^N$, $\mathbf{a}$ is random unit vector as introduced in 
		Section~\ref{sect:alignment}, $c(t)$ is a binary variable drawn 
		from $\{-0.5,0.5\}$ with equal probability and $\zeta_a(t)$ and the
		$\zeta_{dist,i}(t)$ are independent random Gaussian variables with 
		zero mean and unit variance. 
		Hence, $\sigma_a$ simply denotes the standard deviation of each Gaussian
		cluster along the direction of the normal vector $\mathbf{a}$ and was
		set to $\sigma_a = 0.25$. 
		Finally, the set of $\mathbf{v}_{dist,i}$ forms a randomly generated
		orthogonal basis of $N_{dist}$ unit vectors which---just as in 
		Section~\ref{sect:alignment}---are also orthogonal to $\mathbf{a}$.
		The free parameter $s$ parameterized
		the standard deviation along this subspace orthogonal to $\mathbf{a}$.
		As indicated by the time dependence, the Gaussian and binary random
		variables were drawn for each time step. The vectors
		$\mathbf{b}$, $\mathbf{a}$, and $\mathbf{v}_{dist,i}$ were generated
		once before the beginning of a simulation run.
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\centering
			\includegraphics[width=0.75\columnwidth]{illustration_classification}
			\caption{{\bf Input Space for a Simple Linear Classification Task.}
				Two clusters of presynaptic basal activity were generated from 
				multivariate Gaussian distributions. Here, $s$ denotes the standard
				deviation orthogonal to the normal vector of the classification
				hyperplane.}
			\label{fig:illustration_classification}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		For the classification task, we created two output neurons receiving
		the same basal presynaptic input, while the top-down input encoded
		the correct linear classification in a one-hot scheme, that is
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			x_{d,0}(t) &= 1 - \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \\
			x_{d,1}(t) &= \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \; ,
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $\Theta(x)$ is the Heaviside step function.

@@

		The resulting accuracy as a function of $N_{dist}$ and $s$
		is shown in Fig.~\ref{fig:classification_accuracy}.
		Albeit differences were small, the compartment model did
		show a better overall accuracy for the tested parameter range.

@@

		However, for our compartment model, we chose to explicitly set the
		threshold to be the mean value between the high- and low-activity regime
		in our compartment model, i.e.\ $\theta_M = (1+\alpha)/2$. By doing so, LTP was
		preferably induced if both basal and apical input was present at the same
		time. Furthermore, instead of the weight decay term, we chose to keep
		the weight normalization as introduced in \eqref{eq:weight_norm}.
		Obviously, for the point model, the reasoning behind our choice of
		$\theta_M$ did not apply. Still, to provide some level of comparability,
		we also ran simulations with a point model where the sliding threshold was
		calculated as a running average of $y^2$. Furthermore, we did not use
		weight normalization in this case, but chose to use a small weight decay term
		with $\epsilon = 0.1$. The results are shown in 
		Fig.~\ref{fig:classification_accuracy_bcm} (classification task) and 
		Fig.~\ref{fig:classification_accuracy_bcm} (Basal-Apical alignment). 
		While the accuracy of the classification for the point model
		was at most comparable to Hebbian learning, the BCM-like rule for the 
		compartment model significantly increased the accuracy for the tested
		parameter range (compare Fig.~\ref{fig:classification_accuracy}).

		Likewise, the compartment model also significantly benefited from
		the BCM rule in terms of basal-apical alignment, while only marginal
		improvements could be observed for the point model
		(compare Fig.~\ref{fig:corr_dimension_scaling_bcm} 
		with Fig.~\ref{fig:corr_dimension_scaling}).

>>>>

		In a supervised learning scheme, where the top down input
		arriving at the apical compartment acts as the teaching signal,
		the most straight-forward learning rule for the basal synaptic
		weights would be derived from an appropriate loss function,
		such as a mean square error, based on the difference between 
		basal and apical input, i.e.\ $I_{\rm p} - I_{\rm d}$. Theoretical work has 
		investigated possible learning mechanisms
		that could utilize an intracellular error signal
		\citep{Urbanczik_2014,Schiess_2016,Guerguiev_2017}.
		However, a clear experimental
		evidence for a physical quantity encoding such an error 
		is---to our knowledge---yet to be found. 
		On the other hand, Hebbian-type plasticity is extensively
		documented in experiments. Therefore, our work is
		based on the question whether the non-linear interactions
		between basal and apical synaptic input could, when
		combined with a Hebbian plasticity rule, allow a neuron
		to learn to reproduce an apical teaching signal in its
		proximal input.

		In our work, we combine a phenomenological model that generates the output
		firing rate as a function of two streams of synaptic input (subsuming basal and apical input)
		with Hebbian, as well as BCM-like plasticity rules on basal synapses. We hypothesized that 
		this combination of neural activation and plasticity rules would lead to an
		increased correlation between basal and apical input.

		Furthermore, this temporal alignment could potentially facilitate apical inputs to act
		as top-down teaching signals, without the need for an explicit error-driven
		learning rule. Thus, we also test our model in a simple linear 
		supervised classification task and compare it with the performance of a simple
		point neuron equipped with similar plasticity rules.

@@

		In all our numerical experiments, we compare this model with a 
		simple point neuron model that was given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			y(t) = \sigma\left(I_{\rm p}(t) + I_{\rm d}(t) - \theta \right) \; .
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		The apical input $I_{\rm d}$ was generated
		`as is', meaning, it is not dynamically calculated as a superposition
		of multiple presynaptic inputs, but given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			I_{\rm d}(t) = n_d(t) x_d(t) - b_d(t) \; ,
			\label{eq:I_d}
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $n_d(t)$ is a scaling factor, $x_d(t)$ a pre-generated
		discrete time sequence and $b_d(t)$ a bias. Note that $n_d$ and $b_d$ 
		are time dependent since they are subject to adaptation processes 
		that are described in the next section.

		Similarly, $I_{\rm p}(t)$ is given by
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			I_{\rm p}(t) = n_p(t) \sum_{i=1}^{N} x_{p,i}(t) w_i(t) - b_p(t) \; ,
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $N$ is the number of presynaptic afferents, $x_{p,i}(t)$ the
		corresponding sequences and $w_i(t)$ the synaptic efficacies.
		As for $I_{\rm d}(t)$, $n_p(t)$ and $b_p(t)$ is a time dependent
		scaling and bias.

@@

		We implemented a Hebbian plasticity rule for the basal 
		synaptic weights given by the following update equation:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			w_i(t+1) &= w_i(t) + \mu_w \left(x_{p,i}(t) - \tilde{x}_{p,i}(t)\right)
			\left(y(t) - \tilde{y} \right)
			\label{eq:bcm_plast} \\
			\tilde{x}_{p,i}(t+1) &= (1-\mu_{av})\tilde{x}_{p,i}(t) + \mu_{av}x_{p,i}(t+1) \\
			\tilde{y}(t+1) &= (1-\mu_{av})\tilde{y}(t) + \mu_{av}y(t+1)
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Additionally, we use a
		synaptic normalization constraint
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			w_i(t) \rightarrow \frac{w_i(t)}{\lVert \mathbf{w}(t)\rVert}
			\label{eq:weight_norm}
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		in each time step, where $\lVert \mathbf{w}(t)\rVert$ denotes the Euclidean
		norm of the synaptic weight vector. 
%		The threshold $\theta_y$ was
%		chosen such that LTP is present if neural activity is in the
%		high-activity regime (yellow area in Fig.~\ref{fig:comp_model}), while
%		LTD is present in the low-activity regime 
%		(given by $\alpha$, see Fig.~\ref{fig:comp_model}). Hence, we set
%		$\theta_y = (1+\alpha)/2$.

		For comparative reasons, the point neuron model is equipped with the
		same plasticity rule for the proximal weights as \eqref{eq:bcm_plast}. 
%		except that the threshold $\theta_y$ was dynamically adjusted as
%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		\begin{equation}
%			\theta_y(t+1) = \left(1-\mu_{av}\right)\theta_y(t) + \mu_{av} y^2(t) \; , 
%		\end{equation}
%		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		which is an implementation of a running average of the square activity,
%		as used in the classic BCM-rule.

		Additionally, the scaling and bias variables are changing dynamically
		according to the following homeostatic plasticity rules:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			b_p(t+1) &= b_p(t) + \mu_b \left[I_{\rm p}(t) - I_{\rm p}^t\right] \\
			b_d(t+1) &= b_d(t) + \mu_b \left[I_{\rm d}(t) - I_{\rm d}^t\right] \\
			n_p(t+1) &= n_p(t) + \mu_n \left[V_p^t - \left( I_{\rm p}(t) - \tilde{I}_p(t)\right)^2\right] \\
			n_d(t+1) &= n_d(t) + \mu_n \left[V_d^t - \left( I_{\rm d}(t) - \tilde{I}_d(t)\right)^2\right] \\
			\tilde{I}_p(t+1) &= (1-\mu_{av})\tilde{I}_p(t) + \mu_{av} I_{\rm p}(t+1) \\
			\tilde{I}_d(t+1) &= (1-\mu_{av})\tilde{I}_d(t) + \mu_{av} I_{\rm d}(t+1)
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $I_{\rm p}^t$, $I_{\rm d}^t$, $V_p^t$ and $V_p^t$ define targets for the 
		temporal means and variances of $I_{\rm p}$ and $I_{\rm d}$. The dynamic variables 
		$\tilde{I}_p$ and $\tilde{I}_d$ are simply low-pass filtered running 
		averages of $I_{\rm p}$ and $I_{\rm d}$.

@@

		After both $x_{p,i}(t)$ and $x_d(t)$ were generated,
		a simulation was run using all previously described plasticity
		mechanisms until the dynamic variables reached a stationary state.
		After this learning phase, another set of input sequences was generated
		using the previously described protocol and $\rho[I_{\rm p},I_{\rm d}]$ was calculated.
		Note that plasticity was turned off in this phase. This entire procedure
		allowed us to calculate $\rho[I_{\rm p},I_{\rm d}]$ as a function of the distraction
		parameters $s$ and $N_{dist}$. The results for both neuron models 
		is shown in Fig.~\ref{fig:corr_dimension_scaling}. Here, the total number of 
		basal inputs is $N=10$. One can observe a decorrelation transition for 
		both models. However, the compartment model supports a significantly stronger
		distraction in terms of the scaling factor $s$ as compared to the point model.
		This is a first confirmation of our hypothesis that nonlinear interactions
		between basal and apical input could improve learning guided by top-down signals.

@@

		Next, we investigated if the observed differences would also improve
		the performance in an actual supervised learning task.
		For this purpose, we constructed presynaptic basal input as illustrated
		in Fig.~\ref{fig:illustration_classification}.
		Written in vector form,
		each sample from the basal input is generated from the following expression:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			\mathbf{x}_p(t) = \mathbf{b} + \mathbf{a}\left(c(t) + \sigma_a \zeta_a(t) \right) 
			+ s \cdot \sum_{i=1}^{N_{dist}} \zeta_{dist,i}(t) \mathbf{v}_{dist,i} \; .
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $\mathbf{b}$ is a random vector drawn uniformly from
		$(0,1)^N$, $\mathbf{a}$ is random unit vector as introduced in 
		Section~\ref{sect:alignment}, $c(t)$ is a binary variable drawn 
		from $\{-0.5,0.5\}$ with equal probability and $\zeta_a(t)$ and the
		$\zeta_{dist,i}(t)$ are independent random Gaussian variables with 
		zero mean and unit variance. 
		Hence, $\sigma_a$ simply denotes the standard deviation of each Gaussian
		cluster along the direction of the normal vector $\mathbf{a}$ and was
		set to $\sigma_a = 0.25$. 
		Finally, the set of $\mathbf{v}_{dist,i}$ forms a randomly generated
		orthogonal basis of $N_{dist}$ unit vectors which---just as in 
		Section~\ref{sect:alignment}---are also orthogonal to $\mathbf{a}$.
		The free parameter $s$ parameterized
		the standard deviation along this subspace orthogonal to $\mathbf{a}$.
		As indicated by the time dependence, the Gaussian and binary random
		variables are drawn for each time step. The vectors
		$\mathbf{b}$, $\mathbf{a}$, and $\mathbf{v}_{dist,i}$ are generated
		once before the beginning of a simulation run.
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\centering
			\includegraphics[width=0.75\columnwidth]{illustration_classification}
			\caption{{\bf Input Space for a Simple Linear Classification Task.}
				Two clusters of presynaptic basal activity were generated from 
				multivariate Gaussian distributions. Here, $s$ denotes the standard
				deviation orthogonal to the normal vector of the classification
				hyperplane.}
			\label{fig:illustration_classification}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		For the classification task, we set up two output neurons receiving
		the same basal presynaptic input, while the top-down input encoded
		the correct linear classification in a one-hot scheme, that is
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			x_{d,0}(t) &= 1 - \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \\
			x_{d,1}(t) &= \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \; ,
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $\Theta(x)$ is the Heaviside step function.

@@

		The resulting accuracy as a function of $N_{dist}$ and $s$
		is shown in Fig.~\ref{fig:classification_accuracy}.
		Albeit differences are small, the compartment model does
		show a better overall accuracy for the tested parameter range.

@@

		However, for our compartment model, we chose to explicitly set the
		threshold to be the mean value between the high- and low-activity regime
		in our compartment model, i.e.\ $\theta_M = (1+\alpha)/2$. By doing so, LTP is
		preferably induced if both basal and apical input are present at the same
		time. Furthermore, instead of the weight decay term, we chose to keep
		the weight normalization as introduced in \eqref{eq:weight_norm}.
		Obviously, for the point model, the reasoning behind our choice of
		$\theta_M$ did not apply. Still, to provide some level of comparability,
		we also ran simulations with a point model where the sliding threshold was
		calculated as a running average of $y^2$. Furthermore, we did not use
		weight normalization in this case, but chose to use a small weight decay term
		with $\epsilon = 0.1$. The results are shown in 
		Fig.~\ref{fig:classification_accuracy_bcm} (classification task) and 
		Fig.~\ref{fig:classification_accuracy_bcm} (Basal-Apical alignment). 
		While the accuracy of the classification for the point model
		is at most comparable to Hebbian learning, the BCM-like rule for the 
		compartment model significantly increases the accuracy for the tested
		parameter range (compare Fig.~\ref{fig:classification_accuracy}).

		Likewise, the compartment model significantly benefits from
		the BCM rule in terms of basal-apical alignment, while only marginal
		improvements can be observed for the point model
		(compare Fig.~\ref{fig:corr_dimension_scaling_bcm} 
		with Fig.~\ref{fig:corr_dimension_scaling}).
